---
output: rmarkdown::github_document
---

  ```{r, echo = FALSE, purl=FALSE}
knitr::opts_chunk$set(collapse=TRUE,
                      comment="##",
                      fig.retina=2,
                      fig.path = "README_figs/README-")
```

# Portuguese COVID-19 Data

**Last updated:** `r format(Sys.time(), '%a %d %b %Y (%H:%M:%S %Z (%z))')`

## Source

For more information about the data and variables see (https://github.com/dssg-pt/covid19pt-data)

The original data were downloaded from an API provide by VOST (https://covid19-api.vost.pt/Requests/get_entry/)

## User Friendly version.

This repository intends to provide a user friendly CSV version of the Portuguese COVID-19 
data (updated daily - once automated). Download the user friendly version from:

+ Cleaned and user friendly data: [covid19pt_DSSG_Long.csv](https://raw.githubusercontent.com/saghirb/Dados_COVID-19_PT/master/data/covid19pt_DSSG_Long.csv)
+ Original unprocessed data (json to CSV): [covid19pt_DSSG_Orig.csv](https://raw.githubusercontent.com/saghirb/Dados_COVID-19_PT/master/data/covid19pt_DSSG_Orig.csv)


# Example Usage


## Read in the data

Using the `data.table` package to process the data.

```{r readin}
# Load Libraries
library(data.table)
# You can use use: library(here)
suppressPackageStartupMessages(library(here))
library(ggplot2)
library(magrittr)

# Change the ggplot theme.
theme_set(theme_bw())

# Read in data as a data.frame and data.table object.
CV <- fread(here("data", "covid19pt_DSSG_Long.csv"))
str(CV)

# Order data by original variable name and date.
setkeyv(CV, c("origVars", "data"))

# Convert data to a data object in dataset and add a chage from previous day variable.
CV[, data := as.Date(data, format = "%Y-%m-%d")][
  , dayChange := value - shift(value, n=1, fill=NA, type="lag"), by = origVars][
  grepl("^sintomas", origVars), dayChange := NA]
```

## Overall Number of Deaths (daily) by Sex

```{r deathsbySex}
CV[origType=="obitos" & sex %in% c("F", "M") & ageGrp==""] %>%
  ggplot(aes(x=data, y=dayChange, fill=as.factor(sex))) +
  geom_bar(stat = "identity") +
  scale_x_date(date_labels = "%b-%Y") +
  theme(legend.position = "bottom") +
  labs(
    title = "COVID-19 Portugal: Number Daily Deaths",
    x = "Date",
    y = "Number of Deaths",
    fill = "Sex")
```

## Recorded Number of Confirmed COVID-19 Cases by Region

```{r casesbyRegion}
CV[origType=="confirmados" & ageGrp=="" & region!="Portugal"] %>%
  ggplot(., aes(x=data, y=value, colour=region)) +
  geom_line() +
  scale_x_date(date_labels = "%b-%Y") +
  scale_y_log10() +
  theme(legend.position = "bottom") +
  labs(
    title = "COVID-19 Portugal: Number of Confirmed Cases",
    x = "Date",
    y = "Number of Confirmed Cases",
    colour = "Region")
```

## Issue with `dayChange`

Change between days can be negative.

```{r dayChageissues}
CV[dayChange<0][
  , .(data, origVars, value, dayChange)]
```

