---
output: rmarkdown::github_document
---
  
```{r, echo = FALSE, purl=FALSE, message = FALSE}
knitr::opts_chunk$set(collapse=TRUE,
                      comment="##",
                      fig.retina=2,
                      fig.path = "README_figs/README-")
```

![Rendered README](https://github.com/CEAUL/Dados_COVID-19_PT/workflows/Render%20README/badge.svg)

## Daily Portuguese COVID-19 Data

**Last updated: `r format(Sys.time(), "%a %d %b %Y (%H:%M:%S %Z [%z])")`**

```{r dataMetaInfo, echo=FALSE, eval=TRUE, results = "asis"}
suppressPackageStartupMessages(library(here))
dataMetaInfo <- readRDS(here("data", "dataMetaInfo.RData"))
cat(dataMetaInfo)
```

### Download User Friendly Version

+ Download the user friendly data from: **[covid19pt_DSSG_Long.csv](https://raw.githubusercontent.com/CEAUL/Dados_COVID-19_PT/master/data/covid19pt_DSSG_Long.csv)** or use the following direct link in your program:
  + https://raw.githubusercontent.com/CEAUL/Dados_COVID-19_PT/master/data/covid19pt_DSSG_Long.csv
  
+ **Variables**
	+ `data`: Date (Portuguese spelling).
	+ `origVars`: Variable name taken from source data.
	+ `origType`: Orginal variable count type.
	+ `other`: Other types of `origVars`.
	+ `symptoms`: Recorded COVID-19 symptoms.
	+ `sex`: Gender (`F` - Females, `M` - Males, `All` - Females & Males).
	+ `ageGrp`: Age groups in years (`desconhecidos` - unknown).
	+ `ageGrpLower`: Lower limit of age group (useful for sorting).
	+ `ageGrpUpper`: Upper limit of age group.
	+ `region`: Portuguese Regions
	+ `value`: Numeric value.
	+ `valueUnits`: Units for the variable `value`.

<br>

+ Download the original unprocessed data (json to CSV) from: **[covid19pt_DSSG_Orig.csv](https://raw.githubusercontent.com/CEAUL/Dados_COVID-19_PT/master/data/covid19pt_DSSG_Orig.csv)**
  
### Source
  
  For more information about the data and variables see: **https://github.com/dssg-pt/covid19pt-data**
  
  The original data were downloaded from an API provide by VOST **https://covid19-api.vost.pt/Requests/get_entry/**
  
### Summary: Last 10 (available) Days

```{r summary8Days, echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE}
library(data.table)
library(magrittr)
library(knitr)
suppressPackageStartupMessages(library(here)) # library(here)
rDat <- fread(here("data", "covid19pt_DSSG_Orig.csv"))

useVars <- c("Date", "Cases_7_Day_Mean", "Cases", "Active", "Recovered", "Deaths")
cvwd <- rDat[, `:=`(data2 = as.Date(data, format = "%d-%m-%Y"),
                            ativos = as.numeric(ativos))][
      order(data2)][
  , `:=`(confirmados_Avg7 = frollmean(confirmados_novos, 7),
         obitos_novos = obitos - shift(obitos, n=1, fill=NA, type="lag"),
         ativos_novos = ativos - shift(ativos, n=1, fill=NA, type="lag"),
         recuperados_novos = recuperados - shift(recuperados, n=1, fill=NA, type="lag"))][
  , `:=`(Date = format(data2, "%a %d %b %Y"),
         Cases_7_Day_Mean = sprintf("%4.1f", confirmados_Avg7),
         Cases = sprintf("%d (%+d)", confirmados, confirmados_novos),
         Deaths = sprintf("%d (%+d)", obitos, obitos_novos),
         Active = sprintf("%d (%+d)", ativos, ativos_novos),
         Recovered = sprintf("%d (%+d)", recuperados, recuperados_novos))]

cvwd[, tail(.SD, 10)][
    , ..useVars]  %>%
  kable(align = c("rrrrrr"), 
        format = "html",
        table.attr ="class=\"striped\"",
        col.names = c("Date", "Cases 7 Day Mean", "Cases", "Active", "Recovered", "Deaths"),
        caption = "Change from previous day in brackets.")
```

```{r plotNewCases, echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE}
library(ggplot2)
cvwd %>% 
  ggplot(aes(x=data2, y=confirmados_novos)) +
  geom_bar(stat = "identity", fill = "grey75") +
  geom_line(aes(x = data2, y = confirmados_Avg7), group=1, colour = "blue") +
  scale_x_date(date_breaks = "1 months",
               date_labels = "%b-%y",
               limits = c(min(cvwd$data2, na.rm = TRUE), NA)) +
  scale_y_continuous(breaks = seq(0, max(cvwd[, confirmados_novos]), 200)) +
  labs(
    title = "COVID-19 Portugal: Confirmed Daily Cases with 7 Day Rolling Mean",
    x = "",
    y = "Number of Confirmed Cases",
    caption = paste0("Updated on: ", format(Sys.time(), "%a %d %b %Y (%H:%M:%S %Z [%z])"))) +
  theme_bw()
```
  
## Example Usage
  
  
### Read in the data
  
  Using the `data.table` package to process the data.

```{r readinm, message=FALSE}
# Load Libraries
library(data.table)
library(here)

# Read in data as a data.frame and data.table object.
CVPT <- fread(here("data", "covid19pt_DSSG_Long.csv"))
# You can use the direct link:
# CV <- fread("https://raw.githubusercontent.com/CEAUL/Dados_COVID-19_PT/master/data/covid19pt_DSSG_Long.csv")

# Looking at the key variables in the original long dataset.
CVPT[, .(data, origVars, origType, sex, ageGrp, region, value, valueUnits)]

# Order data by original variable name and date.
setkeyv(CVPT, c("origVars", "data"))

# Convert data to a data object in dataset and add a change from previous day variable.
# Added a 7 day rolling average for origVars (except for symptoms). 
# Columns `data` is date in Portuguese.
CV <- CVPT[, data := as.Date(data, format = "%Y-%m-%d")][
  , dailyChange := value - shift(value, n=1, fill=NA, type="lag"), by = origVars][
    grepl("^sintomas", origVars), dailyChange := NA][
  , mean7Day := fifelse(origVars %chin% c("ativos", "confirmados", "obitos", "recuperados"), 
                         frollmean(dailyChange, 7), as.numeric(NA))]

```

### Overall Number of Deaths (daily)

```{r deathsbySex}
library(ggplot2)
library(magrittr)

# Change the ggplot theme.
theme_set(theme_bw())
# Data error prevents by sex plot.
# obMF <- CV[origType=="obitos" & sex %chin% c("M", "F") & ageGrp=="" & region == "Portugal"]
obAll <- CV[origType=="obitos" & sex %chin% c("All") & ageGrp=="" & region == "Portugal"][ 
  , sex := NA]

obAll %>% 
  ggplot(aes(x=data, y=dailyChange)) +
  geom_bar(stat = "identity", fill = "grey75") +
  geom_line(data = obAll, aes(x = data, y = mean7Day), group=1, colour = "brown") +
  scale_x_date(date_breaks = "1 months",
               date_labels = "%b-%y",
               limits = c(min(cvwd$data2, na.rm = TRUE), NA)) +
  theme(legend.position = "bottom") +
  labs(
    title = "COVID-19 Portugal: Number Daily Deaths with 7 Day Rolling Mean",
    x = "",
    y = "Number of Deaths",
    colour = "",
    fill = "",
    caption = paste0("Updated on: ", format(Sys.time(), "%a %d %b %Y (%H:%M:%S %Z [%z])"))
    )
```

### Recorded Number of Confirmed COVID-19 Cases by Age Group

```{r casesbyAgeSex, fig.height=7}
CV[origType=="confirmados" & !(ageGrp %chin% c("", "desconhecidos"))][
  , .(valueFM = sum(value)), .(data, ageGrp)] %>%
  ggplot(., aes(x=data, y=valueFM, colour = ageGrp)) +
  geom_line() +
  scale_x_date(date_breaks = "2 months",
               date_labels = "%b-%y",
               limits = c(min(cvwd$data2, na.rm = TRUE), NA)) +
  scale_y_continuous() +
  theme(legend.position = "bottom") +
  labs(
    title = "COVID-19 Portugal: Number of Confirmed Cases by Age Group",
    x = "",
    y = "Number of Confirmed Cases",
    caption = paste0("Updated on: ", format(Sys.time(), "%a %d %b %Y (%H:%M:%S %Z [%z])")),
    colour = "Age Group")
```

### Recorded Number of Confirmed COVID-19 Cases by Region

```{r casesbyRegion}
CV[origType=="confirmados" & ageGrp=="" & region!="Portugal"] %>%
  ggplot(., aes(x=data, y=value, colour=region)) +
  geom_line() +
  scale_x_date(date_breaks = "1 months",
               date_labels = "%b-%y",
               limits = c(min(cvwd$data2, na.rm = TRUE), NA)) +
  scale_y_log10() +
  theme(legend.position = "bottom") +
  labs(
    title = "COVID-19 Portugal: Number of Confirmed Cases by Region",
    x = "",
    y = "Number of Confirmed Cases",
    caption = paste0("Updated on: ", format(Sys.time(), "%a %d %b %Y (%H:%M:%S %Z [%z])")),
    colour = "Region")
```

<hr>
## Issues & Notes

### Use and interpret with care.

The data are provided as is. Any quality issues or errors in the source data will be 
reflected in the user friend data. 

Please **create an issue** to discuss any errors, issues, requests or improvements.

### Calculated change between days can be negative (`dailyChange`).

```{r dayChageissues}
CV[dailyChange<0 & !(origType %in% c("vigilancia", "internados"))][
  , .(data, origType, origVars, value, dailyChange)]
```

