---
output: rmarkdown::github_document
---
  
```{r, echo = FALSE, purl=FALSE}
knitr::opts_chunk$set(collapse=TRUE,
                      comment="##",
                      fig.retina=2,
                      fig.path = "README_figs/README-")
```

![Rendered README](https://github.com/CEAUL/Dados_COVID-19_PT/workflows/Render%20README/badge.svg)

## Daily Portuguese COVID-19 Data

**Last updated: `r format(Sys.time(), '%a %d %b %Y (%H:%M:%S %Z [%z])**')`

```{r dataMetaInfo, echo=FALSE, eval=TRUE, results = "asis"}
suppressPackageStartupMessages(library(here))
dataMetaInfo <- readRDS(here("data", "dataMetaInfo.RData"))
cat(dataMetaInfo)
```

### Download User Friendly Version

+ Download the user friendly data from: **[covid19pt_DSSG_Long.csv](https://raw.githubusercontent.com/CEAUL/Dados_COVID-19_PT/master/data/covid19pt_DSSG_Long.csv)** or use the following direct link in your program:
  + https://raw.githubusercontent.com/CEAUL/Dados_COVID-19_PT/master/data/covid19pt_DSSG_Long.csv
  
+ **Variables**
	+ `data`: Date (Portuguese spelling).
	+ `origVars`: Variable name taken from source data.
	+ `origType`: Orginal variable count type.
	+ `other`: Other types of `origVars`.
	+ `symptoms`: Recorded COVID-19 symptoms.
	+ `sex`: Gender (`F` - Females, `M` - Males, `All` - Females & Males).
	+ `ageGrp`: Age groups in years (`desconhecidos` - unknown).
	+ `ageGrpLower`: Lower limit of age group (useful for sorting).
	+ `ageGrpUpper`: Upper limit of age group.
	+ `region`: Portuguese Regions
	+ `value`: Numeric value.
	+ `valueUnits`: Units for the variable `value`.

<br>

+ Download the original unprocessed data (json to CSV) from: **[covid19pt_DSSG_Orig.csv](https://raw.githubusercontent.com/CEAUL/Dados_COVID-19_PT/master/data/covid19pt_DSSG_Orig.csv)**
  
### Source
  
  For more information about the data and variables see: **https://github.com/dssg-pt/covid19pt-data**
  
  The original data were downloaded from an API provide by VOST **https://covid19-api.vost.pt/Requests/get_entry/**
  
### Summary: Last 8 Days

```{r summary8Days, echo=FALSE, eval=TRUE}
library(data.table)
library(magrittr)
library(knitr)
suppressPackageStartupMessages(library(here)) # library(here)
rDat <- fread(here("data", "covid19pt_DSSG_Orig.csv"))

useVars <- c("data", "confirmados_novos", "confirmados", "ativos", "recuperados", "obitos_novos", "obitos")
rDat[, `:=`(data = as.Date(data, format = "%d-%m-%Y"),
            obitos_novos = obitos - shift(obitos, n=1, fill=NA, type="lag"))][
  , ..useVars][
  order(data), tail(.SD, 8)] %>% 
  kable()
```

  
## Example Usage
  
  
### Read in the data
  
  Using the `data.table` package to process the data.

```{r readin}
# Load Libraries
library(data.table)
suppressPackageStartupMessages(library(here)) # library(here)

# Read in data as a data.frame and data.table object.
CV <- fread(here("data", "covid19pt_DSSG_Long.csv"))
# You can use the direct link:
# CV <- fread("https://raw.githubusercontent.com/CEAUL/Dados_COVID-19_PT/master/data/covid19pt_DSSG_Long.csv")

# Looking at the data:
tail(CV)

# Order data by original variable name and date.
setkeyv(CV, c("origVars", "data"))

# Convert data to a data object in dataset and add a change from previous day variable.
CV[, data := as.Date(data, format = "%Y-%m-%d")][
  , dailyChange := value - shift(value, n=1, fill=NA, type="lag"), by = origVars][
    grepl("^sintomas", origVars), dailyChange := NA]
```

### Overall Number of Deaths (daily) by Sex

```{r deathsbySex}
library(ggplot2)
library(magrittr)

# Change the ggplot theme.
theme_set(theme_bw())

CV[origType=="obitos" & sex %in% c("F", "M") & ageGrp==""] %>%
  ggplot(aes(x=data, y=dailyChange, fill=as.factor(sex))) +
  geom_bar(stat = "identity") +
  scale_x_date(date_labels = "%b-%Y") +
  theme(legend.position = "bottom") +
  labs(
    title = "COVID-19 Portugal: Number Daily Deaths",
    x = "Date",
    y = "Number of Deaths",
    fill = "Sex")
```

### Recorded Number of Confirmed COVID-19 Cases by Region

```{r casesbyRegion}
CV[origType=="confirmados" & ageGrp=="" & region!="Portugal"] %>%
  ggplot(., aes(x=data, y=value, colour=region)) +
  geom_line() +
  scale_x_date(date_labels = "%b-%Y") +
  scale_y_log10() +
  theme(legend.position = "bottom") +
  labs(
    title = "COVID-19 Portugal: Number of Confirmed Cases",
    x = "Date",
    y = "Number of Confirmed Cases",
    colour = "Region")
```

<hr>
## Issues & Notes

### Use and interpret with care.

The data are provided as is. Any quality issues or errors in the source data will be 
reflected in the user friend data. 

Please **create an issue** to discuss any errors, issues, requests or improvements.

### Calculated change between days can be negative (`dailyChange`).

```{r dayChageissues}
CV[dailyChange<0 & !(origType %in% c("vigilancia", "internados"))][
  , .(data, origType, origVars, value, dailyChange)]
```

